---
title: "THC, 11OH, Cp v time"
output: html_document
date: "2025-11-11"
---

This code was used for Figure 2 to plot the observed plasma concentrations for THC, 11-OH-THC, 11-COOH-THC, and its acyl glucuronide on the control and treated study days. 

The plots are individual points, average line, and standard deviation as shaded gray region. 

The points bloq are not included in the average and are represented as 1/2 lloq and an x in the plots. 

# Set up 

Set the directory to the folder the R markdown is in. 

#### Clear work space 

```{r}

rm(list = ls(all.names = TRUE)); gc()
if (!is.null(dev.list())) graphics.off()
cat("\014")


```

#### Load libraries 

```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(openxlsx)

```

#### Import df and set up to long format

```{r}
df <- read_excel("THC_all_met_Cpt.xlsx", sheet = "All")


# Convert all analyte columns to character so pivot_longer won't fail
df <- df %>%
  mutate(across(
    .cols = c(starts_with("THC_"), starts_with("OH_"),
              starts_with("COOH_"), starts_with("Gluc_")),
    .fns = as.character
  ))


df_long <- df %>%
  pivot_longer(
    cols = c(starts_with("THC_"), starts_with("OH_"),
             starts_with("COOH_"), starts_with("Gluc_")),
    names_to = c("Analyte", "Condition"),
    names_sep = "_",
    values_to = "Concentration_nM"
  ) %>%
  mutate(
    Condition = factor(Condition, levels = c("HC", "HT", "EC", "ET")),
    
    # Recode analyte names for clarity
    Analyte = recode(Analyte,
                     "THC"  = "THC",
                     "OH"   = "11-OH-THC",
                     "COOH" = "11-COOH-THC",
                     "Gluc" = "11-COOH-THC glucuronide"),

    # Order facets: THC > OH > COOH > Gluc
    Analyte = factor(Analyte,
                     levels = c("THC", "11-OH-THC", "11-COOH-THC", "11-COOH-THC glucuronide")),

    Concentration_nM = suppressWarnings(as.numeric(Concentration_nM)),
    Dose_nmol = suppressWarnings(as.numeric(Dose_nmol))
  )

# Replace special cases with NA
df[df == "< 0" | df == "N/A" | df == "NA"] <- NA
df_long[df_long == "< 0" | df_long == "N/A" | df_long == "NA"] <- NA

### create new directory for future 

out_dir <- "Cpt_figures"
if (!dir.exists(out_dir)) dir.create(out_dir)

```

### classification of df

Note that for the LOQs we have determined that quantifiable is 20% within LLOQ. THC = 0.1 nM; OH = 0.1 nM; COOH = 1 nM; Gluc = 3 nM

```{r}

# Define LOQs
loq_thc  <- 0.008
loq_oh   <- 0.080
loq_cooh <- 0.800
loq_gluc <- 2.600

# Pivot long, preserve raw values
df_long_raw <- df %>%
  pivot_longer(
    cols = c(starts_with("THC_"), starts_with("OH_"),
             starts_with("COOH_"), starts_with("Gluc_")),
    names_to = c("Analyte", "SD"),
    names_sep = "_",
    values_to = "raw"
  ) %>%
  mutate(
    raw_chr  = as.character(raw),
    num_val  = suppressWarnings(as.numeric(raw_chr)),

    # Short analyte names for classification logic
    Analyte_short = recode(Analyte,
                           "THC"  = "THC",
                           "OH"   = "OH",
                           "COOH" = "COOH",
                           "Gluc" = "Gluc"),

    # Apply classification
    class = case_when(
      # Missing or coded missing
      raw_chr %in% c("NA", "N/A") | is.na(raw_chr) ~ "BLD",
      str_detect(raw_chr, "<")                    ~ "LT0",

      # THC
      Analyte_short == "THC"  & !is.na(num_val) & num_val > loq_thc  ~ "Quant",
      Analyte_short == "THC"  & !is.na(num_val)                      ~ "BLQ",

      # OH
      Analyte_short == "OH"   & !is.na(num_val) & num_val > loq_oh   ~ "Quant",
      Analyte_short == "OH"   & !is.na(num_val)                      ~ "BLQ",

      # COOH
      Analyte_short == "COOH" & !is.na(num_val) & num_val > loq_cooh ~ "Quant",
      Analyte_short == "COOH" & !is.na(num_val)                      ~ "BLQ",

      # Glucuronide
      Analyte_short == "Gluc" & !is.na(num_val) & num_val > loq_gluc ~ "Quant",
      Analyte_short == "Gluc" & !is.na(num_val)                      ~ "BLQ",

      TRUE ~ "BLD"
    )
  )

# ---- Summary of counts per class ----
summary_table <- df_long_raw %>%
  group_by(Time_h, Analyte_short, SD, class) %>%
  summarise(n = n(), .groups = "drop") %>%
  pivot_wider(
    names_from = class,
    values_from = n,
    values_fill = 0
  ) %>%
  rename(
    n_quant = Quant,
    n_BLQ   = BLQ,
    n_LT0   = LT0,
    n_BLD   = BLD
  ) %>%
  arrange(Analyte_short, SD, Time_h)

View(summary_table)


# Save summary table to Excel
outfile <- file.path(out_dir, "Quant_n_pts_THC_met.xlsx")

write.xlsx(summary_table,
           file = outfile,
           overwrite = TRUE)


```

#### plotting of the quantifable time points 

```{r}

### create directory 
quant_dir <- file.path(out_dir, "Quantifiability")
if (!dir.exists(quant_dir)) dir.create(quant_dir)

### prepare plotting df 
plot_df <- summary_table %>%
  mutate(
    n_BLD_combined = n_BLD + n_LT0
  ) %>%
  select(Time_h, Analyte_short, SD,
         Quantifiable = n_quant,
         BLQ = n_BLQ,
         BLD = n_BLD_combined) %>%
  pivot_longer(
    cols = c(Quantifiable, BLQ, BLD),
    names_to = "Class",
    values_to = "Count"
  ) %>%
  mutate(
    Class = factor(Class,
                   levels = c("Quantifiable", "BLQ", "BLD"))
  )

### plotting funciton 

plot_quant_time <- function(df, analyte, study_day) {

  df_sub <- df %>%
    filter(Analyte_short == analyte,
           SD == study_day)

  ggplot(df_sub, aes(x = Time_h, y = Count,
                     linetype = Class,
                     shape = Class,
                     color = Class)) +

    geom_line(linewidth = 0.9) +
    geom_point(size = 2) +

    # ---- Y axis: increments of 1 ----
    scale_y_continuous(
      limits = c(0, 13),
      breaks = seq(0, 13, 1)
    ) +

    # ---- X axis: increments of 4 ----
    scale_x_continuous(
      breaks = seq(0, max(df_sub$Time_h, na.rm = TRUE), 4)
    ) +

    scale_linetype_manual(values = c(
      Quantifiable = "solid",
      BLQ          = "dashed",
      BLD          = "dotted"
    )) +

    scale_shape_manual(values = c(
      Quantifiable = 16,
      BLQ          = 17,
      BLD          = 15
    )) +

    scale_color_manual(values = c(
      Quantifiable = "black",
      BLQ          = "grey40",
      BLD          = "grey70"
    )) +

    labs(
      x = "Time (h)",
      y = "Number of samples",
      color = "Legend",
      linetype = "Legend",
      shape = "Legend"
    ) +

    theme_bw(base_size = 13) +
    theme(
      panel.grid = element_blank(),
      legend.position = "right"
    )
}


## loope de loop over analytes by study days 

analytes <- unique(plot_df$Analyte_short)
study_days <- unique(plot_df$SD)

for (a in analytes) {
  for (sd in study_days) {

    p <- plot_quant_time(plot_df, analyte = a, study_day = sd)

    fname_base <- paste0("Quant_", a, "_", sd)

    ggsave(
      filename = file.path(quant_dir, paste0(fname_base, ".png")),
      plot = p,
      width = 5,
      height = 4,
      dpi = 300
    )

    ggsave(
      filename = file.path(quant_dir, paste0(fname_base, ".pdf")),
      plot = p,
      width = 5,
      height = 4
    )
  }
}



```


# Averages plots for the hydrocortisone arm control (HC) and treated (HT) study days


```{r}


df_summary <- df_long %>%
  group_by(Analyte, Condition, Time_h) %>%
  summarise(
    mean_conc = mean(Concentration_nM, na.rm = TRUE),
    sd_conc   = sd(Concentration_nM, na.rm = TRUE),
    .groups   = "drop"
  )


```



# Individual Plots 

## HC 

#### THC log 

```{r}

bloq_thc  <- 0.008
bloq_half_thc <- 0.005

# =========================================================================
# Filter HC only

df_THC_HC <- df_long %>%
  filter(Analyte == "THC", Condition == "HC") %>%
  mutate(
    conc_plot = case_when(
      Concentration_nM < bloq_thc ~ bloq_half_thc,
      TRUE ~ Concentration_nM
    ),
    is_blq = Concentration_nM < bloq_thc
  )

# Summary for HC
df_THC_HC_summary <- df_THC_HC %>%
  group_by(Time_h) %>%
  summarise(
    mean_conc = mean(Concentration_nM, na.rm = TRUE),
    sd_conc   = sd(Concentration_nM, na.rm = TRUE),
    .groups = "drop"
  )

plot_THC_HC <- ggplot() +
     
  # Background gridlines (draw first)
  geom_hline(
    yintercept = c(0.001, 0.01, 0.1, 1, 10),
    color = "gray90", linetype = "dashed"
  ) +
  geom_vline(
    xintercept = seq(0, 24, 4),
    color = "gray90", linetype = "dashed"
  ) +

  # SD ribbon
  geom_ribbon(
    data = df_THC_HC_summary,
    aes(
      x = Time_h,
      ymin = pmax(mean_conc - sd_conc, 0.001),
      ymax = mean_conc + sd_conc
    ),
    fill = "gray70", alpha = 0.4
  ) +

  # Mean line
  geom_line(
    data = df_THC_HC_summary,
    aes(x = Time_h, y = mean_conc),
    color = "black", linewidth = 1
  ) +

  # Observed points
  geom_point(
    data = df_THC_HC %>% filter(!is_blq),
    aes(x = Time_h, y = conc_plot),
    color = "black", size = 2
  ) +

  # BLQ points as X
  geom_point(
    data = df_THC_HC %>% filter(is_blq),
    aes(x = Time_h, y = conc_plot),
    shape = 4, size = 3, stroke = 1, color = "black"
  ) +

  # Scales
  scale_y_log10(
    breaks = c(0.001, 0.01, 0.1, 1, 10),
    labels = scales::label_number(), ##this does not set it correctly.. 
    limits = c(0.001, 10)
  ) +
  annotation_logticks(sides = "l") +

  scale_x_continuous(
    breaks = seq(0, 24, 4),
    limits = c(0, 24)
  ) +

  labs(
    x = "Time (h)",
    y = "THC (nM)"
  ) +

  theme_bw(base_size = 13) +
  theme(
    plot.title = element_blank(),
    panel.grid = element_blank()
  )


plot_THC_HC


# =========================================================================
# Filter HT only
df_THC_HT <- df_long %>%
  filter(Analyte == "THC", Condition == "HT") %>%
  mutate(
    conc_plot = case_when(
      Concentration_nM < bloq_thc ~ bloq_half_thc,
      TRUE ~ Concentration_nM
    ),
    is_blq = Concentration_nM < bloq_thc
  )

# Summary for HT
df_THC_HT_summary <- df_THC_HT %>%
  group_by(Time_h) %>%
  summarise(
    mean_conc = mean(Concentration_nM, na.rm = TRUE),
    sd_conc   = sd(Concentration_nM, na.rm = TRUE),
    .groups = "drop"
  )

# HT plot (matching HC style)
plot_THC_HT <- ggplot() +
  
  # Background gridlines
  geom_hline(
    yintercept = c(0.001, 0.01, 0.1, 1, 10),
    color = "gray90", linetype = "dashed"
  ) +
  geom_vline(
    xintercept = seq(0, 24, 4),
    color = "gray90", linetype = "dashed"
  ) +
  
  # SD ribbon (clamped at lower limit)
  geom_ribbon(
    data = df_THC_HT_summary,
    aes(
      x = Time_h,
      ymin = pmax(mean_conc - sd_conc, 0.001),
      ymax = mean_conc + sd_conc
    ),
    fill = "gray70", alpha = 0.4
  ) +
  
  # Mean line
  geom_line(
    data = df_THC_HT_summary,
    aes(x = Time_h, y = mean_conc),
    color = "black", linewidth = 1
  ) +
  
  # Observed points (non-BLQ)
  geom_point(
    data = df_THC_HT %>% filter(!is_blq),
    aes(x = Time_h, y = conc_plot),
    color = "black", size = 2
  ) +
  
  # BLQ points as X
  geom_point(
    data = df_THC_HT %>% filter(is_blq),
    aes(x = Time_h, y = conc_plot),
    shape = 4, size = 3, stroke = 1, color = "black"
  ) +
  
  # Log scale
  scale_y_log10(
    breaks = c(0.001, 0.01, 0.1, 1, 10),
    labels = function(x) format(x, scientific = FALSE, trim = TRUE),
    limits = c(0.001, 10)
  ) +
  annotation_logticks(sides = "l") +
  
  scale_x_continuous(
    breaks = seq(0, 24, 4),
    limits = c(0, 24)
  ) +
  
  labs(
    x = "Time (h)",
    y = "THC (nM)"
  ) +
  
  theme_bw(base_size = 13) +
  theme(
    plot.title = element_blank(),
    panel.grid = element_blank()
  )

plot_THC_HT


# View plots 

plot_THC_HC
plot_THC_HT

```


#### OH log 

```{r}

# --- OH thresholds ---
bloq_oh  <- 0.08
bloq_half_oh <- 0.05
lloq_oh  <- 0.1      # not plotted, just for reference


# HC ============================================================================

# Filter OH HC

df_OH_HC <- df_long %>%
  filter(Analyte == "11-OH-THC", Condition == "HC") %>%
  mutate(
    conc_plot = if_else(Concentration_nM < bloq_oh, bloq_half_oh, Concentration_nM),
    is_blq    = Concentration_nM < bloq_oh
  )

# Use conc_plot for means and SD

df_OH_HC_summary <- df_OH_HC %>%
  group_by(Time_h) %>%
  summarise(
    mean_conc = mean(conc_plot, na.rm = TRUE),
    sd_conc   = sd(conc_plot, na.rm = TRUE),
    .groups = "drop"
  )

plot_OH_HC <- ggplot() +

  # Background
  geom_hline(yintercept = c(0.01, 0.1, 1, 10),
             color = "gray90", linetype = "dashed") +
  geom_vline(xintercept = seq(0, 24, 4),
             color = "gray90", linetype = "dashed") +

  # SD ribbon (clamped to log limits)
  geom_ribbon(
    data = df_OH_HC_summary,
    aes(
      x = Time_h,
      ymin = pmax(mean_conc - sd_conc, 0.01),
      ymax = pmin(mean_conc + sd_conc, 10)
    ),
    fill = "gray70", alpha = 0.4
  ) +

  # Mean line
  geom_line(
    data = df_OH_HC_summary,
    aes(x = Time_h, y = mean_conc),
    color = "black", linewidth = 1
  ) +

  # Observed points
  geom_point(
    data = df_OH_HC %>% filter(!is_blq),
    aes(x = Time_h, y = conc_plot),
    size = 2, color = "black"
  ) +

  # BLQ as X
  geom_point(
    data = df_OH_HC %>% filter(is_blq),
    aes(x = Time_h, y = conc_plot),
    shape = 4, size = 3, stroke = 1, color = "black"
  ) +

  scale_y_log10(
    breaks = c(0.01, 0.1, 1, 10),
    limits = c(0.01, 10)
  ) +
  annotation_logticks(sides = "l") +

  scale_x_continuous(
    breaks = seq(0, 24, 4),
    limits = c(0, 24)
  ) +

  labs(
    x = "Time (h)",
    y = "11-OH-THC (nM)"
  ) +

  theme_bw(base_size = 13) +
  theme(panel.grid = element_blank())


plot_OH_HC


# HT ============================================================================

# Filter OH HT

df_OH_HT <- df_long %>%
  filter(Analyte == "11-OH-THC", Condition == "HT") %>%
  mutate(
    conc_plot = if_else(Concentration_nM < bloq_oh, bloq_half_oh, Concentration_nM),
    is_blq    = Concentration_nM < bloq_oh
  )

df_OH_HT_summary <- df_OH_HT %>%
  group_by(Time_h) %>%
  summarise(
    mean_conc = mean(conc_plot, na.rm = TRUE),
    sd_conc   = sd(conc_plot, na.rm = TRUE),
    .groups = "drop"
  )

plot_OH_HT <- ggplot() +

  geom_hline(yintercept = c(0.01, 0.1, 1, 10),
             color = "gray90", linetype = "dashed") +
  geom_vline(xintercept = seq(0, 24, 4),
             color = "gray90", linetype = "dashed") +

  geom_ribbon(
    data = df_OH_HT_summary,
    aes(
      x = Time_h,
      ymin = pmax(mean_conc - sd_conc, 0.01),
      ymax = pmin(mean_conc + sd_conc, 10)
    ),
    fill = "gray70", alpha = 0.4
  ) +

  geom_line(
    data = df_OH_HT_summary,
    aes(x = Time_h, y = mean_conc),
    color = "black", linewidth = 1
  ) +

  geom_point(
    data = df_OH_HT %>% filter(!is_blq),
    aes(x = Time_h, y = conc_plot),
    size = 2, color = "black"
  ) +

  geom_point(
    data = df_OH_HT %>% filter(is_blq),
    aes(x = Time_h, y = conc_plot),
    shape = 4, size = 3, stroke = 1, color = "black"
  ) +

  scale_y_log10(
    breaks = c(0.01, 0.1, 1, 10),
    limits = c(0.01, 10)
  ) +
  annotation_logticks(sides = "l") +

  scale_x_continuous(
    breaks = seq(0, 24, 4),
    limits = c(0, 24)
  ) +

  labs(
    x = "Time (h)",
    y = "11-OH-THC (nM)"
  ) +

  theme_bw(base_size = 13) +
  theme(panel.grid = element_blank())


plot_OH_HT


# View 

plot_OH_HC
plot_OH_HT

```



#### COOH log 

```{r}

# --- COOH thresholds ---
bloq_cooh      <-  0.8
bloq_half_cooh <-  0.5
lloq_cooh      <-  1


# HC ============================================================================
df_COOH_HC <- df_long %>%
  filter(Analyte == "11-COOH-THC", Condition == "HC") %>%
  mutate(
    conc_plot = if_else(Concentration_nM < bloq_cooh, bloq_half_cooh, Concentration_nM),
    is_blq    = Concentration_nM < bloq_cooh
  )

df_COOH_HC_summary <- df_COOH_HC %>%
  group_by(Time_h) %>%
  summarise(
    mean_conc = mean(conc_plot, na.rm = TRUE),
    sd_conc   = sd(conc_plot, na.rm = TRUE),
    .groups = "drop"
  )

plot_COOH_HC <- ggplot() +
     
  # Background
  geom_hline(
    yintercept = c(0.1, 1, 10, 100),
    color = "gray90", linetype = "dashed"
  ) +
  geom_vline(
    xintercept = seq(0, 24, 4),
    color = "gray90", linetype = "dashed"
  ) +

  # Ribbon
  geom_ribbon(
    data = df_COOH_HC_summary,
    aes(
      x = Time_h,
      ymin = pmax(mean_conc - sd_conc, 0.1),
      ymax = pmin(mean_conc + sd_conc, 200)
    ),
    fill = "gray70", alpha = 0.4
  ) +

  # Mean line
  geom_line(
    data = df_COOH_HC_summary,
    aes(x = Time_h, y = mean_conc),
    color = "black", linewidth = 1
  ) +

  # Points
  geom_point(
    data = df_COOH_HC %>% filter(!is_blq),
    aes(x = Time_h, y = conc_plot),
    size = 2, color = "black"
  ) +
  geom_point(
    data = df_COOH_HC %>% filter(is_blq),
    aes(x = Time_h, y = conc_plot),
    shape = 4, size = 3, stroke = 1, color = "black"
  ) +

  # Y scale
  scale_y_log10(
    breaks = c(0.1, 1, 10, 100),
    limits = c(0.1, 200)
  ) +
  annotation_logticks(sides = "l") +

  # X scale
  scale_x_continuous(
    breaks = seq(0, 24, 4),
    limits = c(0, 24)
  ) +

  labs(
    x = "Time (h)",
    y = "11-COOH-THC (nM)"
  ) +

  theme_bw(base_size = 13) +
  theme(panel.grid = element_blank())

plot_COOH_HC



# HT ============================================================================
df_COOH_HT <- df_long %>%
  filter(Analyte == "11-COOH-THC", Condition == "HT") %>%
  mutate(
    conc_plot = if_else(Concentration_nM < bloq_cooh, bloq_half_cooh, Concentration_nM),
    is_blq    = Concentration_nM < bloq_cooh
  )

df_COOH_HT_summary <- df_COOH_HT %>%
  group_by(Time_h) %>%
  summarise(
    mean_conc = mean(conc_plot, na.rm = TRUE),
    sd_conc   = sd(conc_plot, na.rm = TRUE),
    .groups = "drop"
  )

plot_COOH_HT <- ggplot() +
     
  geom_hline(
    yintercept = c(0.1, 1, 10, 100),
    color = "gray90", linetype = "dashed"
  ) +
  geom_vline(
    xintercept = seq(0, 24, 4),
    color = "gray90", linetype = "dashed"
  ) +

  geom_ribbon(
    data = df_COOH_HT_summary,
    aes(
      x = Time_h,
      ymin = pmax(mean_conc - sd_conc, 0.1),
      ymax = pmin(mean_conc + sd_conc, 200)
    ),
    fill = "gray70", alpha = 0.4
  ) +

  geom_line(
    data = df_COOH_HT_summary,
    aes(x = Time_h, y = mean_conc),
    color = "black", linewidth = 1
  ) +

  geom_point(
    data = df_COOH_HT %>% filter(!is_blq),
    aes(x = Time_h, y = conc_plot),
    size = 2, color = "black"
  ) +
  geom_point(
    data = df_COOH_HT %>% filter(is_blq),
    aes(x = Time_h, y = conc_plot),
    shape = 4, size = 3, stroke = 1, color = "black"
  ) +

  scale_y_log10(
    breaks = c(0.1, 1, 10, 100),
    limits = c(0.1, 200)
  ) +
  annotation_logticks(sides = "l") +

  scale_x_continuous(
    breaks = seq(0, 24, 4),
    limits = c(0, 24)
  ) +

  labs(
    x = "Time (h)",
    y = "11-COOH-THC (nM)"
  ) +

  theme_bw(base_size = 13) +
  theme(panel.grid = element_blank())

plot_COOH_HT


# View 
plot_COOH_HC
plot_COOH_HT



```


#### Gluc log 

```{r}

# --- COOH glucuronide thresholds ---
bloq_gluc      <- 2.6
bloq_half_gluc <- 1.5
lloq_gluc      <- 3


# HC ============================================================================
df_gluc_HC <- df_long %>%
  filter(Analyte == "11-COOH-THC glucuronide", Condition == "HC") %>%
  mutate(
    conc_plot = if_else(Concentration_nM < bloq_gluc, bloq_half_gluc, Concentration_nM),
    is_blq    = Concentration_nM < bloq_gluc
  )

df_gluc_HC_summary <- df_gluc_HC %>%
  group_by(Time_h) %>%
  summarise(
    mean_conc = mean(conc_plot, na.rm = TRUE),
    sd_conc   = sd(conc_plot, na.rm = TRUE),
    .groups = "drop"
  )

plot_gluc_HC <- ggplot() +
     
  # Background
  geom_hline(
    yintercept = c(1, 10, 100),
    color = "gray90", linetype = "dashed"
  ) +
  geom_vline(
    xintercept = seq(0, 24, 4),
    color = "gray90", linetype = "dashed"
  ) +

  # Ribbon
  geom_ribbon(
    data = df_gluc_HC_summary,
    aes(
      x = Time_h,
      ymin = pmax(mean_conc - sd_conc, 1),
      ymax = pmin(mean_conc + sd_conc, 200)
    ),
    fill = "gray70", alpha = 0.4
  ) +

  # Mean line
  geom_line(
    data = df_gluc_HC_summary,
    aes(x = Time_h, y = mean_conc),
    color = "black", linewidth = 1
  ) +

  # Points
  geom_point(
    data = df_gluc_HC %>% filter(!is_blq),
    aes(x = Time_h, y = conc_plot),
    size = 2, color = "black"
  ) +
  geom_point(
    data = df_gluc_HC %>% filter(is_blq),
    aes(x = Time_h, y = conc_plot),
    shape = 4, size = 3, stroke = 1, color = "black"
  ) +

  # Y scale
  scale_y_log10(
    breaks = c(1, 10, 100),
    limits = c(1, 200)
  ) +
  annotation_logticks(sides = "l") +

  # X scale
  scale_x_continuous(
    breaks = seq(0, 24, 4),
    limits = c(0, 24)
  ) +

  labs(
    x = "Time (h)",
    y = "11-COOH-THC glucuronide (nM)"
  ) +

  theme_bw(base_size = 13) +
  theme(panel.grid = element_blank())

plot_gluc_HC



# HT ============================================================================
df_gluc_HT <- df_long %>%
  filter(Analyte == "11-COOH-THC glucuronide", Condition == "HT") %>%
  mutate(
    conc_plot = if_else(Concentration_nM < bloq_gluc, bloq_half_gluc, Concentration_nM),
    is_blq    = Concentration_nM < bloq_gluc
  )

df_gluc_HT_summary <- df_gluc_HT %>%
  group_by(Time_h) %>%
  summarise(
    mean_conc = mean(conc_plot, na.rm = TRUE),
    sd_conc   = sd(conc_plot, na.rm = TRUE),
    .groups = "drop"
  )

plot_gluc_HT <- ggplot() +
     
  geom_hline(
    yintercept = c(1, 10, 100),
    color = "gray90", linetype = "dashed"
  ) +
  geom_vline(
    xintercept = seq(0, 24, 4),
    color = "gray90", linetype = "dashed"
  ) +

  geom_ribbon(
    data = df_gluc_HT_summary,
    aes(
      x = Time_h,
      ymin = pmax(mean_conc - sd_conc, 1),
      ymax = pmin(mean_conc + sd_conc, 200)
    ),
    fill = "gray70", alpha = 0.4
  ) +

  geom_line(
    data = df_gluc_HT_summary,
    aes(x = Time_h, y = mean_conc),
    color = "black", linewidth = 1
  ) +

  geom_point(
    data = df_gluc_HT %>% filter(!is_blq),
    aes(x = Time_h, y = conc_plot),
    size = 2, color = "black"
  ) +
  geom_point(
    data = df_gluc_HT %>% filter(is_blq),
    aes(x = Time_h, y = conc_plot),
    shape = 4, size = 3, stroke = 1, color = "black"
  ) +

  scale_y_log10(
    breaks = c(1, 10, 100),
    limits = c(1, 200)
  ) +
  annotation_logticks(sides = "l") +

  scale_x_continuous(
    breaks = seq(0, 24, 4),
    limits = c(0, 24)
  ) +

  labs(
    x = "Time (h)",
    y = "11-COOH-THC glucuronide (nM)"
  ) +

  theme_bw(base_size = 13) +
  theme(panel.grid = element_blank())

plot_gluc_HT


# View

plot_gluc_HC
plot_gluc_HT



```


#### save plots 

```{r}

save_plot <- function(plot_obj, plot_name) {
  
  # Build file names
  png_file <- file.path(out_dir, paste0(plot_name, ".png"))
  pdf_file <- file.path(out_dir, paste0(plot_name, ".pdf"))
  
  # Save PNG
  ggsave(
    filename = png_file,
    plot = plot_obj,
    width = 5,
    height = 5,
    units = "in",
    dpi = 300
  )
  
  # Save PDF
  ggsave(
    filename = pdf_file,
    plot = plot_obj,
    width = 5,
    height = 5,
    units = "in"
  )
  
  message("Saved: ", plot_name)
}


save_plot(plot_THC_HC,  "plot_THC_HC")
save_plot(plot_THC_HT,  "plot_THC_HT")

save_plot(plot_OH_HC,   "plot_OH_HC")
save_plot(plot_OH_HT,   "plot_OH_HT")

save_plot(plot_COOH_HC, "plot_COOH_HC")
save_plot(plot_COOH_HT, "plot_COOH_HT")

save_plot(plot_gluc_HC, "plot_gluc_HC")
save_plot(plot_gluc_HT, "plot_gluc_HT")


```


