---
title: "Effect Data"
author: "Aurora"
date: "2025-11-22"
output: html_document
---

The aim of this code is to 


# Set up 

Make sure to set to the working directory of the folder 

#### Clear work space 

```{r}

rm(list = ls(all.names = TRUE)); gc()
if (!is.null(dev.list())) graphics.off()
cat("\014")


```

#### Load libraries 

```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(stringr)
library(openxlsx)
library(dplyr)
library(tidyr)
library(purrr)
library(ez)
library(writexl)
library(outliers)
library(ggdist)
library(cowplot)

```


## organize df 

#### load df 

```{r}

df <- read_excel("20251122_PKPD.xlsx", sheet = "NCA")
          # this is the df with the AUClast and Cmax for each study arm 


# loading data for the effect vs time 

HC <- read_excel("20251122_PKPD.xlsx", sheet = "HC")
HT <- read_excel("20251122_PKPD.xlsx", sheet = "HT")


       

```


##### add Cavg to df 

```{r}

df_cavg <- df %>%
filter(Parameter == "AUClast") %>%
mutate(
Parameter = "Cavg",
Value = Value / 24
)

df <- bind_rows(df, df_cavg)


```

##### assign arm and condition 

```{r}

df <- df %>%
     mutate(
          Arm = case_when(
          SD %in% c("HC", "HT") ~ "Hydrocortisone",
     ),
          Condition = case_when(
          SD %in% c("HC") ~ "Control",
          SD %in% c("HT") ~ "Treated"
     )
)


```

#### create long df for paired testing later 

```{r}


df_long <- df %>%
  select(SID, Arm, Effect, Parameter, Condition, Value) %>%
  mutate(logValue = log(Value)) %>%
  drop_na(logValue)


```


# summarize data 

### functions 

```{r}

geo_summary <- function(x) {
x <- x[!is.na(x)]
n <- length(x)

if (n < 2) {
return(tibble(
n = n,
mean = ifelse(n == 0, NA_real_, mean(x)),
geo_mean = ifelse(n == 0, NA_real_, exp(mean(log(x)))),
min = ifelse(n == 0, NA_real_, min(x)),
max = ifelse(n == 0, NA_real_, max(x)),
CI90_low = NA_real_,
CI90_high = NA_real_
))
}

logx <- log(x)
mean_x <- mean(x)
sd_x <- sd(x)
se_x <- sd_x / sqrt(n)
mean_log <- mean(logx)
sd_log   <- sd(logx)
se_log   <- sd_log / sqrt(n)

t_crit <- qt(0.95, df = n - 1)

CI_low <- mean_x - t_crit * se_x 
CI_high <- mean_x + t_crit * se_x

CI_low_log  <- mean_log - t_crit * se_log
CI_high_log <- mean_log + t_crit * se_log

tibble(
n        = n,
mean     = mean_x, 
CI_low   = CI_low, 
CI_high  = CI_high,
geo_mean = exp(mean_log),
min      = min(x),
max      = max(x),
CI90_low_log  = exp(CI_low_log),
CI90_high_log = exp(CI_high_log)
)
}


```

### run smmary 

```{r}

compute_effect_stats <- function(df_group) {

# Control

res_control <- geo_summary(df_group$Value[df_group$Condition == "Control"]) %>%
mutate(Measure = "Control")

# Treated

res_treated <- geo_summary(df_group$Value[df_group$Condition == "Treated"]) %>%
mutate(Measure = "Treated")

# Treated / Control (paired)

tc_df <- df_group %>%
select(SID, Condition, Value) %>%
pivot_wider(names_from = Condition, values_from = Value) %>%
drop_na(Control, Treated) %>%
mutate(TC = Treated / Control)

res_tc <- geo_summary(tc_df$TC) %>%
mutate(Measure = "TC_ratio")

bind_rows(res_control, res_treated, res_tc)
}


```

### compute 

```{r}

summary_stats <- df %>%
group_by(Arm, Effect, Parameter) %>%
group_modify(~ compute_effect_stats(.x)) %>%
ungroup() %>%
relocate(Arm, Effect, Parameter, Measure)


```

# paired t-test, grubbs, and FDR 

### function 

```{r}


run_paired_tests <- function(df_long_arm) {
  
  effects_here <- unique(df_long_arm$Effect)
  results_all <- list()
  
  for (eff in effects_here) {
    
    df_e <- df_long_arm %>% filter(Effect == eff)
    params <- unique(df_e$Parameter)
    
    res_list <- list()
    
    for (p in params) {
      
      df_p <- df_e %>% filter(Parameter == p)
      
      df_wide <- df_p %>%
        select(SID, Condition, logValue) %>%
        pivot_wider(names_from = Condition, values_from = logValue) %>%
        drop_na(Control, Treated)
      
      N_used <- nrow(df_wide)
      if (N_used < 2) next
      
      # Paired t-test
      tt <- t.test(df_wide$Treated, df_wide$Control, paired = TRUE)
      
      # Grubbs test (no deletion)
      Gc <- tryCatch(grubbs.test(df_wide$Control), error = function(e) NA)
      Gt <- tryCatch(grubbs.test(df_wide$Treated), error = function(e) NA)
      
      res_list[[p]] <- tibble(
        Arm = unique(df_p$Arm),
        Effect = eff,
        Parameter = p,
        N = N_used,
        mean_log_Control = mean(df_wide$Control),
        mean_log_Treated = mean(df_wide$Treated),
        mean_log_diff = mean(df_wide$Treated - df_wide$Control),
        t_stat = tt$statistic,
        p_raw = tt$p.value,
        
        # Grubbs
        G_control = ifelse(is.list(Gc), Gc$statistic, NA),
        p_G_control = ifelse(is.list(Gc), Gc$p.value, NA),
        
        G_treated = ifelse(is.list(Gt), Gt$statistic, NA),
        p_G_treated = ifelse(is.list(Gt), Gt$p.value, NA)
      )
    }
    
    results_e <- bind_rows(res_list)
    
    # Apply FDR within Arm × Effect
    results_e <- results_e %>%
      mutate(p_FDR = p.adjust(p_raw, method = "BH"))
    
    results_all[[eff]] <- results_e
  }
  
  bind_rows(results_all)
}


```

#### run 

```{r}

df_HC <- df_long %>% filter(Arm == "Hydrocortisone")


paired_HC <- run_paired_tests(df_HC)

```


# save 

### output

```{r}

output_dir <- "Effect_PD_Data"
if (!dir.exists(output_dir)) {
dir.create(output_dir)
}

# figures 

fig_dir <- file.path(output_dir, "Figures")
if (!dir.exists(fig_dir)) dir.create(fig_dir)

```

#### summary stats 

```{r}

effects <- unique(summary_stats$Effect)

wb <- createWorkbook()

for (eff in effects) {
df_eff <- summary_stats %>%
filter(Effect == eff) %>%
arrange(Arm, Parameter, Measure)

addWorksheet(wb, eff)
writeData(wb, eff, df_eff)
}

# Save file

saveWorkbook(wb, file.path(output_dir, "Summary.xlsx"), overwrite = TRUE)


```

# create wide df 

```{r}

# CREATE WIDE DATA FOR EXPORT (CONTROL, TREATED, TC COLUMNS)


# Compute Control, Treated, and TC for each SID × Effect × Parameter
wide_base <- df %>%
  select(SID, Arm, Condition, Effect, Parameter, Value) %>%
  pivot_wider(
    names_from = Condition,
    values_from = Value
  ) %>%
  mutate(
    TC = ifelse(!is.na(Control) & !is.na(Treated),
                Treated / Control,
                NA_real_)
  )

# Pivot wider so that HR_Cmax_Control, HR_Cmax_Treated, HR_Cmax_TC, etc. become columns
wide_data <- wide_base %>%
  mutate(
    col_Control = paste0(Effect, "_", Parameter, "_Control"),
    col_Treated = paste0(Effect, "_", Parameter, "_Treated"),
    col_TC      = paste0(Effect, "_", Parameter, "_TC")
  ) %>%
  select(
    SID, Arm,
    col_Control, Control,
    col_Treated, Treated,
    col_TC, TC
  ) %>%
  pivot_longer(
    cols = c(Control, Treated, TC),
    names_to = "Measure",
    values_to = "Value"
  ) %>%
  mutate(
    FinalCol = case_when(
      Measure == "Control" ~ col_Control,
      Measure == "Treated" ~ col_Treated,
      Measure == "TC"      ~ col_TC
    )
  ) %>%
  select(SID, Arm, FinalCol, Value) %>%
  pivot_wider(
    names_from = FinalCol,
    values_from = Value
  ) %>%
  arrange(Arm, SID)

#Split into Hydrocortisone 
wide_hc <- wide_data %>% filter(Arm == "Hydrocortisone") %>% select(-Arm)


# SAVE OUTPUT EXCEL FILE

output_dir <- "PD_Effect_Data"
if (!dir.exists(output_dir)) dir.create(output_dir)

file_out <- file.path(output_dir, "data_wide.xlsx")

wb2 <- createWorkbook()

addWorksheet(wb2, "Hydrocortisone")
writeData(wb2, "Hydrocortisone", wide_hc)


saveWorkbook(wb2, file_out, overwrite = TRUE)


```


#### paired t tests

```{r}

file_paired <- file.path(output_dir, "Paired_TTests.xlsx")

wb_paired <- createWorkbook()

addWorksheet(wb_paired, "Hydrocortisone")
writeData(wb_paired, "Hydrocortisone", paired_HC)

saveWorkbook(wb_paired, file_paired, overwrite = TRUE)


```



# PD Effects  

### saving plots 

```{r}

save_plot <- function(plot_object, filename, width = 6, height = 4, dpi = 300) {
  ggsave(
    filename = file.path(fig_dir, filename),
    plot = plot_object,
    width = width,
    height = height,
    dpi = dpi
  )
}


```


### HR and BP

```{r}

make_time_plots <- function(df, time_col, value_col, arm_label, effect_label,
                            y_label = "Value",
                            file_prefix = "Plot") {

  # ---- REMOVE NA ROWS FOR THE VALUE COLUMN ----
  df_clean <- df %>%
    filter(!is.na(.data[[value_col]]),
           !is.na(.data[[time_col]]))   # also ensure Time isn't NA

  # Create summary for mean and SD
  df_summary <- df_clean %>%
    group_by(.data[[time_col]]) %>%
    summarise(
      mean_val = mean(.data[[value_col]], na.rm = TRUE),
      sd_val   = sd(.data[[value_col]], na.rm = TRUE),
      .groups = "drop"
    )

  # -------------------------
  # Plot 1: 0–12 hours
  # -------------------------
  plot_0_12 <- ggplot() +
    geom_ribbon(
      data = df_summary,
      aes(x = .data[[time_col]],
          ymin = mean_val - sd_val,
          ymax = mean_val + sd_val),
      fill = "gray70", alpha = 0.4
    ) +
    geom_line(
      data = df_summary,
      aes(x = .data[[time_col]], y = mean_val),
      color = "black", linewidth = 1
    ) +
    geom_point(
      data = df_clean,
      aes(x = .data[[time_col]], y = .data[[value_col]]),
      color = "black", size = 2
    ) +
    scale_x_continuous(
      limits = c(0, 12),
      breaks = seq(0, 12, 4)
    ) +
    labs(
      x = "Time (h)",
      y = y_label
    ) +
    theme_bw(base_size = 13) +
    theme(
      panel.grid = element_blank(),
      plot.title = element_blank()
    )

  # Save 0–12 h plot
  ggsave(
    filename = file.path(fig_dir,
                         paste0(file_prefix, "_", arm_label, "_", effect_label, "_0_12h.png")),
    plot = plot_0_12,
    width = 6, height = 4, dpi = 300
  )


  # -------------------------
  # Plot 2: 0–24 hours
  # -------------------------
  plot_0_24 <- ggplot() +
    geom_ribbon(
      data = df_summary,
      aes(x = .data[[time_col]],
          ymin = mean_val - sd_val,
          ymax = mean_val + sd_val),
      fill = "gray70", alpha = 0.4
    ) +
    geom_line(
      data = df_summary,
      aes(x = .data[[time_col]], y = mean_val),
      color = "black", linewidth = 1
    ) +
    geom_point(
      data = df_clean,
      aes(x = .data[[time_col]], y = .data[[value_col]]),
      color = "black", size = 2
    ) +
    scale_x_continuous(
      limits = c(0, 24),
      breaks = seq(0, 24, 4)
    ) +
    labs(
      x = "Time (h)",
      y = y_label
    ) +
    theme_bw(base_size = 13) +
    theme(
      panel.grid = element_blank(),
      plot.title = element_blank()
    )

  # Save 0–24 h plot
  ggsave(
    filename = file.path(fig_dir,
                         paste0(file_prefix, "_", arm_label, "_", effect_label, "_0_24h.png")),
    plot = plot_0_24,
    width = 6, height = 4, dpi = 300
  )


  # ---- RETURN PLOTS FOR HTML DISPLAY ----
  list(
    plot_0_12 = plot_0_12,
    plot_0_24 = plot_0_24
  )
}



```

#### HC 

```{r}

# --- HC: HEART RATE ---
p <- make_time_plots(
  df = HC,
  time_col = "Time",
  value_col = "HR",
  arm_label = "HC",
  effect_label = "HR",
  y_label = "Heart Rate (bpm)",
  file_prefix = "HR"
)

p$plot_0_12 <- p$plot_0_12 + coord_cartesian(ylim = c(0, 120))
p$plot_0_24 <- p$plot_0_24 + coord_cartesian(ylim = c(0, 120))

ggsave(
  filename = file.path(fig_dir, "HR_HC_HR_0_12h.png"),
  plot = p$plot_0_12,
  width = 6, height = 4, dpi = 300
)

ggsave(
  filename = file.path(fig_dir, "HR_HC_HR_0_24h.png"),
  plot = p$plot_0_24,
  width = 6, height = 4, dpi = 300
)


# --- HC: SYSTOLIC BP ---
p <- make_time_plots(
  df = HC,
  time_col = "Time",
  value_col = "Systolic_BP",
  arm_label = "HC",
  effect_label = "Systolic",
  y_label = "Systolic BP (mmHg)",
  file_prefix = "SystolicBP"
)

p$plot_0_12 <- p$plot_0_12 + coord_cartesian(ylim = c(0, 140))
p$plot_0_24 <- p$plot_0_24 + coord_cartesian(ylim = c(0, 140))

ggsave(file.path(fig_dir, "SystolicBP_HC_Systolic_0_12h.png"),
       p$plot_0_12, width = 6, height = 4, dpi = 300)

ggsave(file.path(fig_dir, "SystolicBP_HC_Systolic_0_24h.png"),
       p$plot_0_24, width = 6, height = 4, dpi = 300)


# --- HC: DIASTOLIC BP ---
p <- make_time_plots(
  df = HC,
  time_col = "Time",
  value_col = "Diastolic_BP",
  arm_label = "HC",
  effect_label = "Diastolic",
  y_label = "Diastolic BP (mmHg)",
  file_prefix = "DiastolicBP"
)

p$plot_0_12 <- p$plot_0_12 + coord_cartesian(ylim = c(0, 100))
p$plot_0_24 <- p$plot_0_24 + coord_cartesian(ylim = c(0, 100))

ggsave(file.path(fig_dir, "DiastolicBP_HC_Diastolic_0_12h.png"),
       p$plot_0_12, width = 6, height = 4, dpi = 300)

ggsave(file.path(fig_dir, "DiastolicBP_HC_Diastolic_0_24h.png"),
       p$plot_0_24, width = 6, height = 4, dpi = 300)


# --- HC: PULSE PRESSURE ---
p <- make_time_plots(
  df = HC,
  time_col = "Time",
  value_col = "Pulse_pressure",
  arm_label = "HC",
  effect_label = "PP",
  y_label = "Pulse Pressure (mmHg)",
  file_prefix = "PP"
)

p$plot_0_12 <- p$plot_0_12 + coord_cartesian(ylim = c(0, 80))
p$plot_0_24 <- p$plot_0_24 + coord_cartesian(ylim = c(0, 80))

ggsave(file.path(fig_dir, "PP_HC_PP_0_12h.png"),
       p$plot_0_12, width = 6, height = 4, dpi = 300)

ggsave(file.path(fig_dir, "PP_HC_PP_0_24h.png"),
       p$plot_0_24, width = 6, height = 4, dpi = 300)




```


#### HT

```{r}

# --- HT: HEART RATE ---
p <- make_time_plots(
  df = HT,
  time_col = "Time",
  value_col = "HR",
  arm_label = "HT",
  effect_label = "HR",
  y_label = "Heart Rate (bpm)",
  file_prefix = "HR"
)

p$plot_0_12 <- p$plot_0_12 + coord_cartesian(ylim = c(0, 120))
p$plot_0_24 <- p$plot_0_24 + coord_cartesian(ylim = c(0, 120))

ggsave(file.path(fig_dir, "HR_HT_HR_0_12h.png"),
       p$plot_0_12, width = 6, height = 4, dpi = 300)


ggsave(file.path(fig_dir, "HR_HT_HR_0_24h.png"),
       p$plot_0_24, width = 6, height = 4, dpi = 300)


# --- HT: SYSTOLIC BP ---
p <- make_time_plots(
  df = HT,
  time_col = "Time",
  value_col = "Systolic_BP",
  arm_label = "HT",
  effect_label = "Systolic",
  y_label = "Systolic BP (mmHg)",
  file_prefix = "SystolicBP"
)

p$plot_0_12 <- p$plot_0_12 + coord_cartesian(ylim = c(0, 140))
p$plot_0_24 <- p$plot_0_24 + coord_cartesian(ylim = c(0, 140))

ggsave(file.path(fig_dir, "SystolicBP_HT_Systolic_0_12h.png"),
       p$plot_0_12, width = 6, height = 4, dpi = 300)

ggsave(file.path(fig_dir, "SystolicBP_HT_Systolic_0_24h.png"),
       p$plot_0_24, width = 6, height = 4, dpi = 300)


# --- HT: DIASTOLIC BP ---
p <- make_time_plots(
  df = HT,
  time_col = "Time",
  value_col = "Diastolic_BP",
  arm_label = "HT",
  effect_label = "Diastolic",
  y_label = "Diastolic BP (mmHg)",
  file_prefix = "DiastolicBP"
)

p$plot_0_12 <- p$plot_0_12 + coord_cartesian(ylim = c(0, 100))
p$plot_0_24 <- p$plot_0_24 + coord_cartesian(ylim = c(0, 100))

ggsave(file.path(fig_dir, "DiastolicBP_HT_Diastolic_0_12h.png"),
       p$plot_0_12, width = 6, height = 4, dpi = 300)

ggsave(file.path(fig_dir, "DiastolicBP_HT_Diastolic_0_24h.png"),
       p$plot_0_24, width = 6, height = 4, dpi = 300)


# --- HT: PULSE PRESSURE ---
p <- make_time_plots(
  df = HT,
  time_col = "Time",
  value_col = "Pulse_pressure",
  arm_label = "HT",
  effect_label = "PP",
  y_label = "Pulse Pressure (mmHg)",
  file_prefix = "PP"
)

p$plot_0_12 <- p$plot_0_12 + coord_cartesian(ylim = c(0, 80))
p$plot_0_24 <- p$plot_0_24 + coord_cartesian(ylim = c(0, 80))

ggsave(file.path(fig_dir, "PP_HT_PP_0_12h.png"),
       p$plot_0_12, width = 6, height = 4, dpi = 300)

ggsave(file.path(fig_dir, "PP_HT_PP_0_24h.png"),
       p$plot_0_24, width = 6, height = 4, dpi = 300)



```



# Rain cloud for VAS 

### new directory 

```{r}

fig_dir_raincloud_xcont <- file.path(output_dir, "Figures_raincloud_x_cont")
if (!dir.exists(fig_dir_raincloud_xcont)) {
  dir.create(fig_dir_raincloud_xcont)
}


```

### continous raincloud function

```{r}

make_vas_raincloud_continuous <- function(df, vas_var, time_points,
                                          y_label,
                                          x_label = "Time (h)",
                                          cloud_width = 0.45,
                                          cloud_offset = 0.25) {

  df_plot <- df %>%
    filter(Time %in% time_points,
           !is.na(.data[[vas_var]]))

  ggplot(df_plot, aes(x = Time, y = .data[[vas_var]])) +

    # ---- CLOUD (half violin, pushed right) ----
    stat_halfeye(
      aes(group = factor(Time)),
      side = "right",
      width = cloud_width,
      adjust = 2,
      justification = -cloud_offset,
      alpha = 0.45,
      fill = "darkblue"
    ) +

    # ---- BOX (slimmer) ----
    geom_boxplot(
      aes(group = factor(Time)),
      width = 0.12,
      outlier.shape = NA,
      fill = "gray",
      colour = "darkblue",
      linewidth = 0.45,
      alpha = 0.45
    ) +

    # ---- POINTS ----
    geom_jitter(
      aes(group = factor(Time)),
      width = 0.06,
      height = 0,
      size = 1.6,
      shape = 21,
      stroke = 0.35,
      fill = "gray",
      colour = "black",
      alpha = 0.9
    ) +

    # ---- AXES ----
    scale_x_continuous(
      breaks = c(0, 4, 8, 12, 24),
      minor_breaks = time_points,
      name = x_label
    ) +

    scale_y_continuous(
      limits = c(0.8, 6.2),
      breaks = 1:6
    ) +

    labs(y = y_label) +
    theme_cowplot(font_size = 13) +
    theme(
      axis.ticks.x.minor = element_line(color = "grey80", linewidth = 0.3)
    )
}


```

### study arm and time window 
```{r}

study_days <- list(
  HC = HC,
  HT = HT
)

time_0_12 <- sort(unique(HC$Time[HC$Time >= 0 & HC$Time <= 12]))
time_0_24 <- sort(unique(HC$Time[HC$Time >= 0 & HC$Time <= 24]))


time_sets <- list(
  `0_12h` = time_0_12,
  `0_24h` = time_0_24
)

```

### continous rain cloud 

```{r}

for (day in names(study_days)) {

  df_day <- study_days[[day]]

  for (vas in vas_vars) {

    label <- vas_labels[[vas]]

    for (tset in names(time_sets)) {

      time_points <- time_sets[[tset]]

      p <- make_vas_raincloud_continuous(
        df = df_day,
        vas_var = vas,
        time_points = time_points,
        y_label = label
      )

      filename <- paste0(
        "VAS_", vas, "_", day, "_", tset, "_xcont.png"
      )

      ggsave(
        filename = file.path(fig_dir_raincloud_xcont, filename),
        plot = p,
        width = 6,
        height = 4,
        dpi = 300
      )
    }
  }
}

```

